Chord: A Versatile Platform for Program Analysis

### 2011-03-12 main/doc/acknowledgments.tex
more documentation

\begin{itemize}
\item
\href{http://joeq.sourceforge.net/}{Joeq}, a Java compiler framework
\item
\href{http://www.csg.is.titech.ac.jp/~chiba/javassist/}{Javassist},
a Java bytecode manipulation tool
\item
\href{ttp://bddbddb.sourceforge.net/}{bddbddb}, a BDD-based Datalog solver
\end{itemize}

Chord additionally relies on the following open-source tools and libraries:
\begin{itemize}
\item
\href{http://ant-contrib.sourceforge.net/}{Ant-Contrib},
a collection of useful Ant tasks  
\item
\href{http://buddy.sourceforge.net/}{BuDDy}, a BDD library
\item
\href{http://trove4j.sourceforge.net/}{GNU Trove},
a primitive collections library for Java
\item
\href{http://www.java2html.com/}{Java2HTML} and
\href{http://www.java2html.de/}{Java2Html}, Java to HTML tools
\item
\href{http://saxon.sourceforge.net/}{Saxon}, an XSLT processor
\end{itemize}

>>>>

\begin{itemize}
\item
\xlink{Joeq}{http://joeq.sourceforge.net/}, a Java compiler framework
\item
\xlink{Javassist}{http://www.csg.is.titech.ac.jp/~chiba/javassist/},
a Java bytecode manipulation tool
\item
\xlink{bddbddb}{http://bddbddb.sourceforge.net/}, a BDD-based Datalog solver
\end{itemize}

Chord additionally relies on the following open-source tools and libraries:
\begin{itemize}
\item
\xlink{Ant-Contrib}{http://ant-contrib.sourceforge.net/},
a collection of useful Ant tasks  
\item
\xlink{BuDDy}{http://buddy.sourceforge.net/}, a BDD library
\item
\xlink{GNU Trove}{http://trove4j.sourceforge.net/},
a primitive collections library for Java
\item
\xlink{Java2HTML}{http://www.java2html.com/} and
\xlink{Java2Html}{http://www.java2html.de/}, Java-to-HTML tools
\item
\xlink{Saxon}{http://saxon.sourceforge.net/}, an XSLT processor
\end{itemize}

### 2011-03-12 main/doc/computing_scope.tex
more documentation

\section{Computing program scope}
\label{sec:building-scope}

Chord computes the scope (i.e., reachable methods) of the given
program either if property \code{chord.build.scope} is set to {\tt
  true} or if some other task (e.g., a program analysis specified via
property \code{chord.run.analyses}) demands it by calling method
\code{chord.program.Program.g()}.

Users can set the following properties to control scope computation:

\begin{itemize}
\item \code{chord.main.class}
\item \code{chord.class.path}
\item \code{chord.reuse.scope}
\item \code{chord.methods.file}
\item \code{chord.reflect.file}
\item \code{chord.scope.kind}
\item \code{chord.ch.kind}
\item \code{chord.reflect.kind}
\item \code{chord.std.scope.exclude}
\item \code{chord.ext.scope.exclude}
\item \code{chord.scope.exclude}
\end{itemize}

@@

The possible legal values of property \code{chord.scope.kind} are
[\code{dynamic}$|$\code{rta}$|$\code{cha}].  In each case, Chord at
least expects properties \code{chord.main.class} and
\code{chord.class.path} to be set.

\begin{itemize}
\item
The {\tt dynamic} value instructs Chord to compute program scope
dynamically, by running the program and observing using JVMTI the
classes that are loaded at run-time.  The number of times the program
is run and the command-line arguments to be supplied to the program in
each run is specified by properties \code{chord.run.ids} and
\code{chord.args.<id>} for each run ID {\tt <id>}.  By default, the
program is run only once, using run ID {\tt 0}, and without any
command-line arguments.  Only classes loaded in some run are regarded
as reachable but {\it all} methods of each loaded class are regarded
as reachable regardless of whether they were invoked in the run.  The
rationale behind this decision is to both reduce run-time overhead of
JVMTI and to increase the predictive power of program analyses
performed using the computed program scope.

\item
The {\tt rta} value instructs Chord to compute program scope
statically using Rapid Type Analysis (RTA).
RTA is an iterative fixed-point algorithm.  It maintains a set of
reachable methods $M$.  The initial iteration starts by assuming that
only the main method in the main class is reachable (Chord also
handles class initializer methods but we ignore them here for brevity;
we also ignore the set of reachable classes maintained besides the set
of reachable methods).  All object allocation sites $H$ contained in
methods in $M$ are deemed reachable (i.e., control-flow within method
bodies is ignored).  Whenever a dynamically-dispatching method call
site (i.e., an invokevirtual or invokeinterface site) with receiver of
static type $t$ is encountered in a method in $M$, only subtypes of
$t$ whose objects are allocated at some site in $H$ are considered to
determine the possible target methods, and each such target method is
added to $M$.  The process terminates when no more methods can be
added.

%RTA is a relatively inexpensive and precise algorithm in practice.
%Its key shortcoming is that it makes no attempt to resolve
%reflection, which is rampant in real-world Java programs, and can
%therefore be unsound (i.e., underestimate the set of reachable
%classes and methods).  The next option attempts to overcome this
%problem.  \item The \code{rta_reflect} value instructs Chord to
%compute program scope statically using Rapid Type Analysis and,
%moreover, to resolve a common reflection
%pattern: \begin{quote} \begin{verbatim} String s = ...; Class c =
%Class.forName(s); Object o = c.newInstance(); T t = (T)
%o; \end{verbatim} \end{quote} This analysis is identical to RTA
%except that it additionally inspects every cast statement in the
%program, such as the last statement in the above snippet, and queries
%the class hierarchy to find all concrete classes that subclass
%\code{T} (if \code{T} is a class) or that implement \code{T} (if
%\code{T} is an interface).  Chord allows users to control which
%classes are included in the class hierarchy (see Section
%\ref{sec:cha}).
\item
The {\tt cha} value instructs Chord to compute program scope
statically using Class Hierarchy Analysis (CHA).
The key difference between CHA and RTA is that for invokevirtual and
invokeinterface sites with receiver of static type $t$, CHA considers
{\it all} subtypes of $t$ in the class hierarchy to determine the
possible target methods, whereas RTA restricts them to types of
objects allocated in methods deemed reachable so far.  As a result,
CHA is highly imprecise in practice, and also expensive since it
grossly overestimates the set of reachable classes and methods.
Nevertheless, Chord allows users to control which classes are
included in the class hierarchy (see \ref{sec:cha}) and thereby control the
precision and cost of CHA.

@@

\begin{enumerate}
\item
The boot classpath, specified by property \code{sun.boot.class.path}.
\item
The library extensions classpath, comprising all jar files in
directory \code{[java.home]/lib/dir/}.
\item
Chord's classpath, specified by property
\code{chord.main.class.path}.  This includes the classpath of any
user-defined Java analyses, specified by user-defined property
\code{chord.java.analysis.path}, which is empty by default.
\item
The classpath of the given program, specified by user-defined property
\code{chord.class.path}, which is empty by default.
\end{enumerate}

All classes in the entire classpath (resulting from items 1--4 above)
are included in the class hierarchy with the following exceptions:
\begin{itemize}
\item
Duplicate classes, i.e., classes with the same name occurring in more
than one classpath element; in this case, all occurrences except the
first are excluded.
\item
Any class whose name's prefix is specified in the value of property
\code{chord.scope.exclude}.
\item
If property \code{chord.ch.kind} has value {\tt dynamic}, then
Chord runs the given program and observes the set of all classes the
JVM loads; any class not in this set is excluded.  By default,
property \code{chord.ch.kind} has value {\tt static}.
\item
If the superclass of a class C is missing or if an interface
implemented/extended by a class/interface C is missing, where
``missing" means that it is either not in the classpath resulting from
items 1--4 above or it is excluded by one of these rules, then C
itself is excluded.  Note that this rule is recursive, e.g., if C has
superclass B which in turn has superclass A, and A is missing, then
both B and C are excluded.
\end{itemize}

@@

\begin{itemize}
\item
Property \code{chord.std.scope.exclude} conventionally specifies
classes to be excluded from the scope of {\it all} programs to be
analyzed, e.g., classes in the JDK standard library.  Its default
value is \code{[chord.main.class.path]}, i.e., the classes comprising
Chord itself.
\item
Property \code{chord.ext.scope.exclude} conventionally specifies
classes to be excluded from the scope of specific programs to be
analyzed.  Its default value is the empty list.
\item
Property \code{chord.scope.exclude} specifies the final list of
classes to be excluded from scope.  Its default value is simply the
concatenation of the values of properties
\code{chord.std.scope.exclude} and \code{chord.ext.scope.exclude}.
\end{itemize}

>>>>

\section{Computing Analysis Scope}
\label{sec:computing-scope}

Chord computes the analysis scope (i.e., reachable methods) of the given
program either if property \code{chord.build.scope} is set to {\tt
  true} or if some other task (e.g., a program analysis specified via
property \code{chord.run.analyses}) demands it by calling method
\code{chord.program.Program.g()}.
The following sections describe Chord's analysis scope 
computation in more detail.
%Users can set the following properties to control scope computation:
%
%\begin{itemize}
%\item \code{chord.main.class}
%\item \code{chord.class.path}
%\item \code{chord.reuse.scope}
%\item \code{chord.methods.file}
%\item \code{chord.reflect.file}
%\item \code{chord.scope.kind}
%\item \code{chord.ch.kind}
%\item \code{chord.reflect.kind}
%\item \code{chord.std.scope.exclude}
%\item \code{chord.ext.scope.exclude}
%\item \code{chord.scope.exclude}
%\end{itemize}
%
%The meaning of these properties is explained on demand in

@@

The possible legal values of property \code{chord.scope.kind} are
[\code{dynamic}$|$\code{rta}$|$\code{cha}].  In each case, Chord at
least expects properties \code{chord.main.class} and
\code{chord.class.path} to be set to the fully-qualified name of the
program's main class (e.g., \code{com.example.Main}) and the
program's application classpath, respectively.

\begin{itemize}
\item
The {\tt dynamic} value instructs Chord to compute analysis scope
dynamically, by running the program and observing the
classes that are loaded at run-time.  The number of times the program
is run and the command-line arguments to be supplied to the program in
each run is specified by properties \code{chord.run.ids} and
\code{chord.args.<id>} for each run ID {\tt <id>}.  By default, the
program is run only once, using run ID {\tt 0}, and without any
command-line arguments.  Only classes loaded in some run are regarded
as reachable but {\it all} methods of each loaded class are regarded
as reachable regardless of whether they were invoked in the run.  The
rationale behind this decision is to both reduce the run-time instrumentation
overhead and increase the predictive power of program analyses
performed using the computed analysis scope.

\item
The {\tt rta} value instructs Chord to compute analysis scope
statically using Rapid Type Analysis (RTA).
RTA is an iterative fixed-point algorithm.  It maintains a set of
reachable methods $M$.  The initial iteration starts by assuming that
only the main method in the main class is reachable (Chord also
handles class initializer methods but we ignore them here for brevity;
we also ignore the set of reachable classes maintained besides the set
of reachable methods).  All object allocation sites $H$ contained in
methods in $M$ are deemed reachable (i.e., control-flow within method
bodies is ignored).  Whenever a dynamically-dispatching method call
site (i.e., an invokevirtual or invokeinterface site) with receiver of
static type $t$ is encountered in a method in $M$, only subtypes of
$t$ whose objects are allocated at some site in $H$ are considered to
determine the possible target methods, and each such target method is
added to $M$.  The process terminates when no more methods can be
added.

%RTA is a relatively inexpensive and precise algorithm in practice.
%Its key shortcoming is that it makes no attempt to resolve
%reflection, which is rampant in real-world Java programs, and can
%therefore be unsound (i.e., underestimate the set of reachable
%classes and methods).  The next option attempts to overcome this
%problem.  \item The \code{rta_reflect} value instructs Chord to
%compute analysis scope statically using Rapid Type Analysis and,
%moreover, to resolve a common reflection
%pattern: \begin{quote} \begin{verbatim} String s = ...; Class c =
%Class.forName(s); Object o = c.newInstance(); T t = (T)
%o; \end{verbatim} \end{quote} This analysis is identical to RTA
%except that it additionally inspects every cast statement in the
%program, such as the last statement in the above snippet, and queries
%the class hierarchy to find all concrete classes that subclass
%\code{T} (if \code{T} is a class) or that implement \code{T} (if
%\code{T} is an interface).  Chord allows users to control which
%classes are included in the class hierarchy (see Section
%\ref{sec:cha}).
\item
The {\tt cha} value instructs Chord to compute analysis scope
statically using Class Hierarchy Analysis (CHA).
The key difference between CHA and RTA is that for invokevirtual and
invokeinterface sites with receiver of static type $t$, CHA considers
{\it all} subtypes of $t$ in the class hierarchy to determine the
possible target methods, whereas RTA restricts them to types of
objects allocated in methods deemed reachable so far.  As a result,
CHA is highly imprecise in practice, and also expensive since it
grossly overestimates the set of reachable classes and methods.
Nevertheless, Chord allows users to control which classes are
included in the class hierarchy (see Section \ref{sec:cha}) and thereby control the
precision and cost of CHA.

@@

\begin{enumerate}
\item
The boot classpath, specified by property \code{sun.boot.class.path}.
\item
The library extensions classpath, comprising all jar files in
directory \code{[java.home]/lib/dir/}.
\item
The application classpath of the given program, specified by property \code{chord.class.path},
which is empty by default.
\end{enumerate}

All classes in the entire classpath (resulting from items 1--3 above)
are included in the class hierarchy with the following exceptions:
\begin{itemize}
\item
Duplicate classes, i.e., classes with the same name occurring in more
than one classpath element; in this case, all occurrences except the
first are excluded.
\item
Any class whose name's prefix is specified in the value of property
\code{chord.scope.exclude}.
\item
If property \code{chord.ch.kind} has value {\tt dynamic}, then
Chord runs the given program and observes the set of all classes the
JVM loads; any class not in this set is excluded.  By default,
property \code{chord.ch.kind} has value {\tt static}.
\item
If the superclass of a class C is missing or if an interface
implemented/extended by a class/interface C is missing, where
``missing" means that it is either not in the classpath resulting from
items 1--3 above or it is excluded by one of these rules, then C
itself is excluded.  Note that this rule is recursive, e.g., if C has
superclass B which in turn has superclass A, and A is missing, then
both B and C are excluded.
\end{itemize}

@@

\begin{itemize}
\item
Property \code{chord.std.scope.exclude} conventionally specifies
classes to be excluded from the scope of {\it all} programs to be
analyzed, e.g., classes in the JDK standard library.  Its default
value is the empty list.
\item
Property \code{chord.ext.scope.exclude} conventionally specifies
classes to be excluded from the scope of specific programs to be
analyzed.  Its default value is the empty list.
\item
Property \code{chord.scope.exclude} specifies the final list of
classes to be excluded from scope.  Its default value is
\code{[chord.std.scope.exclude]:[chord.ext.scope.exclude]}.
\end{itemize}

### 2011-03-12 main/doc/deadlock.tex
more documentation

\noindent where {\tt <...>} denotes the absolute or relative path of a directory
containing a file named \code{chord.properties} which defines
properties \code{chord.main.class}, \code{chord.class.path}, and
\code{chord.src.path}.  See Section \ref{sec:chord-sysprops} for the
meaning of these properties.

>>>>

where {\tt <...>} denotes the absolute or relative path of a directory
containing a file named \code{chord.properties} which defines
properties \code{chord.main.class}, \code{chord.class.path}, and
\code{chord.src.path}.  See Section \ref{sec:chord-sysprops} for the
meaning of these properties.

### 2011-03-12 main/doc/setting_up.tex
more documentation

\noindent The above structure is typical: the program's Java source
files are under {\tt src/}, its class files are under {\tt classes/},
the source and jar files of the libraries used by the program are
under \code{lib/src/} and \code{lib/jar/}, respectively.

@@

\noindent where \code{<...>} denotes the absolute or relative path of the
directory containing the above \code{example/} directory.
Property \code{chord.work.dir} specifies the working directory during Chord's execution.
Chord searches for a file named \code{chord.properties} under that working directory, and
loads upfront all properties defined in that file, if it exists.  Alternatively, these properties
may be defined on the command-line, using the \code{-D<key>=<value>} format.
A sample \code{chord.properties} file for the above program is as follows:

@@

In the above case, Chord does not do much beyond loading the above
\code{chord.properties} file.  For Chord to do something interesting,
you need to set additional properties, either on the above command
line or in the \code{chord.properties} file, specifying the task(s)
Chord must perform.  All tasks are summarized in Section
\ref{sec:func-props}.  The two most common tasks, described
next, are computing the analysis scope of the given program (Section
\ref{sec:building-scope}) and running analyses on the given program
(Section \ref{sec:running-analysis}).

>>>>

The above structure is typical: the program's Java source
files are under {\tt src/}, its class files are under {\tt classes/},
the source and jar files of the libraries used by the program are
under \code{lib/src/} and \code{lib/jar/}, respectively.

@@

where \code{<...>} denotes the absolute or relative path of the
directory containing the above \code{example/} directory.
Property \code{chord.work.dir} specifies the working directory during Chord's execution.
Chord searches for a file named \code{chord.properties} under that working directory, and
loads upfront all properties defined in that file, if it exists.  Alternatively, these properties
may be defined on the command-line, using the \code{-D<key>=<value>} format.
A sample \code{chord.properties} file for the above program is as follows:

@@

In the above case, Chord does not do much beyond loading the above
\code{chord.properties} file.  For Chord to do something interesting,
you need to set additional properties, either on the above command
line or in the \code{chord.properties} file, specifying the task(s)
Chord must perform.  All tasks are summarized under ``Functionality Properties"
in Section \ref{sec:chord-sysprops}.  The two most common tasks, described
next, are computing the analysis scope of the given program (Section
\ref{sec:computing-scope}) and running analyses on the given program
(Section \ref{sec:running-analysis}).

### 2011-03-12 main/doc/troubleshooting.tex
more documentation

\noindent where {\tt <sys-props>} denotes a whitespace-separated list of zero of more property definitions
of the form {\tt -D<key>=<value>} (see Section \ref{sec:chord-sysprops} for an exhaustive listing of all
properties recognized by \Chord).

>>>>

where {\tt <sys-props>} denotes a whitespace-separated list of zero of more property definitions
of the form {\tt -D<key>=<value>} (see Section \ref{sec:chord-sysprops} for an exhaustive listing of all
properties recognized by \Chord).

### 2011-03-12 main/doc/writing_dynamic_analysis.tex
more documentation

You can change the default values of various properties for configuring your dynamic analysis;
see Sections \ref{sec:scope-props} and \ref{sec:instr-props}.
For instance:
\begin{itemize}
\item
You can set property \code{chord.scope.kind} to {\tt dynamic} so that the program scope
is computed dynamically (i.e., by running the program) instead of statically.
\item
You can exclude certain classes (e.g., JDK classes) from being instrumented by setting
properties \code{chord.std.scope.exclude}, \code{chord.ext.scope.exclude}, and
\code{chord.scope.exclude}.
\item
You can choose between online (i.e. load-time) and offline bytecode instrumentation by
setting property \code{chord.instr.kind} to {\tt online} or {\tt offline}.
\item
You can require the event-generating and event-handling JVMs to be one and the same (by setting
property \code{chord.trace.kind} to {\tt none}),
or to be separate (by setting property {\tt chord.trace.kind} to {\tt full} or {\tt pipe}, depending upon
whether you want the two JVMs to exchange events by a regular file or a POSIX pipe, respectively).
Using a single JVM can cause correctness/performance issues if event-handling Java code itself is instrumented
(e.g., say the event-handling code uses class \code{java.util.ArrayList} which is not excluded from program scope).
Using separate JVMs prevents such issues since the event-handling JVM runs uninstrumented bytecode (only
the event-generating JVM runs instrumented bytecode).
If a regular file is used to exchange events between the two JVMs, then the JVMs run serially:
the event-generating JVM first runs to completion, dumps the entire dynamic trace to the regular file,
and then the event-handling JVM processes the dynamic trace.
If a POSIX pipe is used to exchange events between the two JVMs, then the JVMs run in lockstep.
Obviously, a pipe is more efficient for very long traces, but it 
not portable (e.g., it does not currently work on Windows/Cygwin), and the traces cannot be reused
across Chord runs (see the following item).
\item
You can reuse dynamic traces from a previous Chord run by setting property
\code{chord.reuse.traces} to {\tt true}.  In this case, you must also set property
{\tt chord.trace.kind} to {\tt full}.
\item
You can set property \code{chord.dynamic.haltonerr} to {\tt false} to prevent Chord from terminating
even if the program on which dynamic analysis is being performed crashes.
\end{itemize}

@@

Compile the analysis by placing the directory containing class \code{MyDynamicAnalysis} created
above in the path defined by property \code{chord.java.analysis.path}.

@@

>>>>

You can change the default values of various properties for configuring your dynamic analysis;
see ``Analysis Scope Properties" and ``Instrumentation Properties" in Section \ref{sec:chord-sysprops}.
For instance:
\begin{itemize}
\item
You can set property \code{chord.scope.kind} to {\tt dynamic} so that the program scope
is computed dynamically (i.e., by running the program) instead of statically.
\item
You can exclude certain classes (e.g., JDK classes) from being instrumented by setting
properties \code{chord.std.scope.exclude}, \code{chord.ext.scope.exclude}, and
\code{chord.scope.exclude}.
\item
You can choose between online (i.e. load-time) and offline bytecode instrumentation by
setting property \code{chord.instr.kind} to {\tt online} or {\tt offline}.
\item
You can require the event-generating and event-handling JVMs to be one and the same (by setting
property \code{chord.trace.kind} to {\tt none}),
or to be separate (by setting property {\tt chord.trace.kind} to {\tt full} or {\tt pipe}, depending upon
whether you want the two JVMs to exchange events by a regular file or a POSIX pipe, respectively).
Using a single JVM can cause correctness/performance issues if event-handling Java code itself is instrumented
(e.g., say the event-handling code uses class \code{java.util.ArrayList} which is not excluded from program scope).
Using separate JVMs prevents such issues since the event-handling JVM runs uninstrumented bytecode (only
the event-generating JVM runs instrumented bytecode).
If a regular file is used to exchange events between the two JVMs, then the JVMs run serially:
the event-generating JVM first runs to completion, dumps the entire dynamic trace to the regular file,
and then the event-handling JVM processes the dynamic trace.
If a POSIX pipe is used to exchange events between the two JVMs, then the JVMs run in lockstep.
Obviously, a pipe is more efficient for very long traces, but it 
not portable (e.g., it does not currently work on Windows/Cygwin), and the traces cannot be reused
across Chord runs (see the following item).
\item
You can reuse dynamic traces from a previous Chord run by setting property
\code{chord.reuse.traces} to {\tt true}.  In this case, you must also set property
{\tt chord.trace.kind} to {\tt full}.
\item
You can set property \code{chord.dynamic.haltonerr} to {\tt false} to prevent Chord from terminating
even if the program on which dynamic analysis is being performed crashes.
\end{itemize}

@@

Compile the analysis by placing the directory containing class \code{MyDynamicAnalysis} created
above in the path defined by property \code{chord.ext.java.analysis.path}.

@@

{\bf Note:} The IBM J9 JVM on Linux is highly recommended if you intend to use Chord for
dynamic program analysis, as it allows you to instrument the entire JDK; using any other
platform will likely require excluding large parts of the JDK from being instrumented.
Additionally, if you intend to use online (load-time) bytecode instrumentation in
your dynamic program analysis, then you will need JDK 6 or higher, since this functionality
requires the \code{java.lang.instrument} API with class retransformation support (the latter
support is available only in JDK 6 and higher).
